
%\part*{Randomized Algorithm : Lecture 1A}

\section{Introduction to Randomized algorithms}
We consider here an undirected graph with no weight on its edges. Let us call the graph $G(V,E)$. A cut
in the graph is a partition of the vertices into two non empty sets $S$ and $\overline{S}$. We define 
the weight of a cut as the number of edges that pass between any vertex $u\in S$ and another vertex $v \in \overline{S}$.
The problem of finding a min-cut is partitioning the vertices of the graph in two sets such that the
number of edges passing between the two partitions is minimum. The problem can be formally stated as follows:
\begin{center}
\textbf{Input} : An unweighted undirected graph $G(V,E)$.\\
\textbf{Output} : A cut $(S,\overline{S})$ with $S \subseteq V$ of minimumm size. 
\end{center}

\noindent\textbf{Deterministic Algorithm} : The deterministic algorithm for finding a mincut relies
on the invocation of the maximum flow algorithm a polynomial number of times. The best known time
bound for such an algorithm is $O(mnlog(n^{2}/m))$.


\noindent\textbf{Randomized Algorithm} : Let us define a multigraph to be a graph with possibly multiple
parallel edges between vertices. Given an edge $(x,y)$ in a multigraph $G=(V,E)$, contraction of
edge $(x,y)$ corresponds to replacing the vertices $x$ and $y$ by another vertex $z$, and for each 
$v\notin \{x,y\}$ replacing any edge $(x,v)$ or $(y,v)$ by the edge $(z,v)$; keeping the remaining
graph same. This would create some multiple edges.\\

\begin{lemma}
 Let $G^{\prime}$ be the graph obtained by contracting an edge in $G$ then the size of the mincut of
$G^{\prime}$ is at least equal to the size of the mincut of $G$. The equality holds if the edge 
contracted does not belong to one of the mincut of G.
\end{lemma}
\begin{proof}
 Clearly, by definition, by contracting an edge, we loose at most the edges that are parallel to it. 
Thus, if the edge contracted were in the min-cut in $G$, the value of the min-cut in $G^{\prime}$ is strictly
greater than or equal to the min-cut in $G$ as it changes the partition of the vertices. In other cases, 
contracting the edge would have no effect on the min-cut as the edge contracted is within one of the 
component of the mincut. Thus, no edge in the cut is affected while contracting the edge.
\end{proof}

As a consequence, we have the following lemma,\\
\begin{lemma}
 Any cut of $G^{\prime}$ is a cut of $G$.
\end{lemma}
Now, we can specify the algorithm, which would contract a randomly picked edges of a graph, and 
then show that the algorithm succeeds to produce the mincut of a graph with high probability.

\noindent \textbf{Algorithm:}

\indent \hspace{5pt} As long as $|V| \leq 3$\\
\indent \indent \hspace{5pt} Draw a random edge $e$ in $G$.\\
\indent \indent \hspace{5pt} Contract e.\\
\indent \hspace{5pt} Output the cut. \\

\noindent \textbf{Analysis of the Algorithm:}\\
As long as we don't contract an edge from the min-cut, the output will be correct. Assuming the size
of the min-cut is $k$ and the number of edges is $m$.\\
\begin{equation*}
  \begin{array}{ll}
	 Pr\{\mbox{Algorithm fails on the first step}\}& = \dfrac{\mbox{Size of min-cut}}{\mbox{edges in graph}}\\
	 & = \dfrac{k}{m}\\
  \end{array}
\end{equation*}

\begin{lemma}
 In an $n$-vertex multigraph with min-cut value $k$, no vertex has degree smaller than $k$. Further 
the total number of edges in the graph satisfies $m\geq nk/2$.
\end{lemma}
\begin{proof}
 By contradiction let us suppose that there is a vertex $v$ with degree less than $k$. Then the cut 
$(\{v\},V\backslash \{v\})$ will be less than $k$, which contradicts the minimality of min-cut.
\end{proof}

\noindent Thus, we have that, Pr\{Algorithm fails on the first step\} = $\dfrac{2}{n}$

\begin{theorem}
 The probability that the algorithm outputs a proper min-cut is at least $2/n^{2}$.
\end{theorem}
\begin{proof}
 We consider the $i^{th}$ step of the algorithm. We say the algorithm succeeds in a step if it contracts
as edge which is not part of a min-cut. We have,\\
Pr\{success at the $i^{th}$ contraction $\arrowvert$ success at all the previous steps\}\\
\begin{equation*}
 \begin{array}{ll}
  &= 1 - Pr\{\mbox{failure at $i^{th}$ contraction $\arrowvert$ no failure before $i^{th}$ step}\}\\
  &\geq 1- \dfrac{2}{n + 1 -i} = \dfrac{n-1-i}{n+1-i}\\ 
 \end{array}
\end{equation*}

\noindent We know, Pr\{min-cut as output\} $\geq$ Pr\{ $\forall step$ i=1 to n-2, there was success\}

\noindent Let, $A_{t}$ denote the event that step $t$ is a success.

\noindent Therefore,
\begin{equation*}
 \begin{array}{ll}
    Pr\{\mbox{min-cut as output}\}&\geq Pr\{A_{1} \wedge A_{2} \wedge \cdots \wedge A_{n-2}\}\\
    &= Pr\{A_{n-2}\arrowvert A_{1} \wedge \cdots \wedge A_{n-3}\}.Pr\{A_{n-3}\arrowvert A_{1} \wedge\cdots \wedge A_{n-4}\}\\
    &  \cdots Pr\{A_{2}\arrowvert A_{1}\} . Pr\{A_{1}\}\\
  \end{array}
\end{equation*}

\noindent(For conditional probability, we have
Pr\{$A\arrowvert B$\}=$\dfrac{Pr\{A\wedge B\}}{Pr\{B\}}$)

\begin{equation*}
 \begin{array}{ll}
    Pr\{\mbox{min-cut as output}\}&\geq \displaystyle\prod_{i=1}^{n-2} \dfrac{(n-1-i)}{(n+1-i)}\\
    &= \dfrac{2}{n(n-1)} \geq \dfrac{2}{n^{2}}\\
  \end{array}
\end{equation*}

\end{proof}

We would like to increase the probability of the algorithm of outputting a min-cut to $1- \dfrac{1}{n^{c}}$
for any constant $c > 0$. For this we run the algorithm for k times and output the minimum of the k cuts
produced by the algorithm. We select k depending on the c. To calculate the value of k, we note that,\\

Pr\{output is not a min-cut in all the k runs\} $\leq \big(1-\dfrac{2}{n^{2}}\big)^k$\\
Thus, Pr\{output is a min-cut after k runs\} $\geq 1-\big(1- \dfrac{2}{n^{2}}\big)^k$\\

Clearly, to make this probability to the desired value we need to choose the value of $k$ such that:
$\big(1- \dfrac{2}{n^{2}}\big)^k = \dfrac{1}{n^{c}}$\\
From, $\big(1- \dfrac{2}{n^{2}}\big)^k$ we have,\\
\begin{equation*}
 exp(k log(1-\dfrac{2}{n^{2}})) \leq exp(-\dfrac{2k}{n^{2}})
\end{equation*}
If we choose $k=\dfrac{c}{2}n^{2}\log n$ then we get the required probability.

\begin{theorem}
 The algorithm runs in time $O(n^{4}\log n)$ on any n-vertex multigraph.
\end{theorem}

\begin{proof}
 This may seem surprising that the algorithm doesn't depend on the number of edges in the graph, as 
in a multigraph the number of edges are not bounded by ${n}\choose{2}$. But, each group of multiedge
can be replaced by weight on a single edge with weights specifying the multiplicity of the edge. Thus,
the number of edges are bounded by ${n}\choose{2}$. And thus the algorithm may be implemented in a way
to terminate in $O(n^{2})$. Repeating the algorithm $k$ times takes time $O(n^{2}k)$, which according
to the value of $k$ is $O(n^{4}\log n)$
\end{proof}
